---
title: "What the Package Does"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{What the Package Does}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MATH4753F25EmilyPeterson)
```


# Introduction

This package contains functions and data sets used in the Applied Statistical Methods class. They are useful in learning about and understanding statistics in various ways.

Below are explanations of each function and data set contained in the package, as well as examples of how to use them.


## birthday

This function calculates the probability of two or more people in a given group size having the same birthday. This is a classic statistics problem demonstrating counting rules.

To use the function, simply input the size of the group. It will output the probability.

```{r}
birthday(20)
```


## DDT

This is a data set from "Statistics for Engineering and the Sciences" Sixth Edition, by William M. Mendenhall and Terry L. Sincich from CRC Press, 2016. It records attributes of fish found in the Tennessee River and its tributaries, including length, weight, and DDT contamination levels.

```{r}
head(ddt)
```


## Fire

This is a data set from "Statistics for Engineering and the Sciences" Sixth Edition, by William M. Mendenhall and Terry L. Sincich from CRC Press, 2016. It describes the distance and damage caused by fires.


```{r}
head(fire)
```


## myboot2

This function uses the bootstrapping method to estimate a confidence interval for the statistic function given to it. It also requires a starting sample to resample from and the number of resamples to preform. It creates a histogram of the data and interval, and returns a list with helpful data.

```{r}
myboot2(10000, x = c(1, 1, 2, 2, 3, 3, 3, 3, 2, 4, 5, 5), fun = "mean", alpha = 0.1, xlab = "mean", cx = 1.5)
```


## mycltu

This function demonstrates the Cnetral Limit Theorem for the uniform distribution. It takes in the number and size of the samples and the minimum and maximum allowed values. It returns a histogram showing that as the number of samples increases, the distribution follows the normal distribution.

```{r}
mycltu(n = 3, iter = 10000, a = 2, b = 5)
```


## myfirst

This function takes in a vector and returns a graph of the values where y is the square of the input values. It also returns a list of the same.


```{r}
myfirstfunction(1:10)
```


## myncurve

This function takes the mean and standard deviation of a distribution and displays the distribution plot with the shaded area of the probability of the given value. It also returns all of this as a list.


```{r}
myncurve(2, 3, 1)
```


## mysample

This function creates barplots showing the distribution of numbers picked from 1-10.

```{r}
mysample(n = 1000, iter = 4)
```


## mysum

This function returns the summary of the given object. It uses summary() and was our first practice in making functions.

```{r}
summary(1:10)
```


## ntickets

This function calculates the number of plane tickets to be sold to maximize profits. It takes in the number of seats on the plane, the probability of actually overbooking, and the probability of someone showing up when they've bought a ticket. It displays the graph for the calculations in the discrete and continuous cases, and returns a list of the important values.

```{r}
ntickets(400, 0.02, 0.95)
```


## scatterhist

This function takes in two vectors of values to plot as a scatterplot of x vs y. It also displays a marginal histogram.

```{r}
scatterhist(x = 1:10, y = 11:20)
```


## shinymle

This function creates a shiny app that was generated by ChatGPT, attempting to demonstrate MLE for several different univariate distributions with adjustable variables. The app is launched using "shinymle()", but it doesn't actually function correctly.


